Charged pion emission from neutron-rich heavy ion collisions for studies on the symmetry energy

### 2020-08-09 chapters_altFormat/ch_experiment.tex
Added Betty's edits

A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector. TPCs have been employed in gases, liquid, and even solid detectors \cite{starTPC,arTPC,eosTPC}. In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this Thesis, along with the rest of the experimental setup.

@@

The pad plane is perpendicular to the electric field of the field cage and is divided into a two dimensional grid of pads. Two of the 3 coordinates are determined from the 2-dimensional charge distribution corresponding to the projection of the track onto the pad-plane. The third dimension comes from extrapolating the electrons back in time, utilizing the known constant drift velocity $v_d$. The distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the drift time information obtained from combining the time of the event trigger with arrival time of the signal from the  electrons in the electronics. The radius of curvature of the track on the pad plane is related to the magnetic rigidity, and therefore the momentum of each track. The energy loss deposited $\langle dE/dx\rangle$ is proportional to the image charge induced on the pads on the pad-plane. With these two observations, particles can be uniquely identified by their mass and charge since there is a unique correlation between the rigidity and $\langle dE/dx\rangle$  lines. This correlation allows the particle type to be identified as discussed later in Section~\ref{sec:pid}. In this chapter we will provide more details about the particle detection within the \spirit TPC  that is relevant to the reactions studied in this thesis.

@@

The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) was developed to measure pions and other light charge particles produced in collisions of radioactive rare isotope beams with stable targets in fixed target experiments.  The TPC is built on an aluminum angle iron skeleton with thin aluminum sheet walls in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC. The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}. The dipole gap limited the vertical space of the TPC to around \SI{75}{\centi\metre}. More detail and specifications of the SAMURAI dipole magnet are given in \cite{samurai}.

The exploded drawing shown in Fig.~\ref{fig:tpcExplode} illustrates these major internal components of the  of the \spirit TPC. A target mechanism allowed for up to 5 fixed targets to be mounted on the target ladder situated just upstream of the field cage, where changing the targets could be performed outside of the magnet. Reaction products from nuclear collisions exit the target and enter the field cage of the TPC. The field cage contains the detector gas and defines the constant electric field. It is rigidly mounted to the large aluminum top plate of the TPC gas containent vessel. The field cage is electrically isolated from the top plate by a lexan top perimeter ring with o-rings to provide as gas seal. The pad plane and wire plane structures are  mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate. Several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a precise laser measurement \cite{jon}. Holes in the top plate allowed signals from the individual charge sensitive pads on the pad plane to be transmitted via cables to the electronics.

@@

\begin{comment}
\subsection{Enclosure}
The skeleton of the enclosure is composed of a rigid rectangular aluminum angle-iron frame. To this frame, six rectangular sides are attached. The downstream window and the two side windows are constructed of a aluminum window frames with a 1 mm  aluminum sheet metal window. These thin sheet metal windows allow charged particles and neutrons to exit the TPC without significant energy loss or scattering. This allows for a trigger to be created by placing detectors outside of  the side windows and downstream window of the TPC. The enclosure itself is made to be gas tight with no leakage from the outside or from the interior of the field cage. This is to allow for the possibility to choosing one insulation gas outside of the field cage and a different counter gas inside the field cage. In the first campaign of experiments, however, we used the same gas in the field cage and enclosure volume.  
\end{comment}

@@

The field cage was constructed from several panels of two-layer printed circuit board (PCBs). The front of the field cage was made of two PCBs and each side was constructed from three PCBs. These boards had copper strips etched onto the inside and outside of the board which when all the boards were assembed, formed copper rings which were used to set up the uniform electric field, as will be described later. The array of circuit boards were glued and supported by Lexan pieces. We did not use the common PCB substrate material FR4, which contains a Bromine impregnated epoxy which can outgas Bromine, which can absorb the secondary electrons from the track, thereby degrading the signal \cite{tpcAging}. Instead, a halogen free material, Cryogenic-G10, was used for the board material.  The downstream wall mainly consists of a large, thin exit window. The window  is constructed of   \SI{10}{\micro\metre} Kapton upon which are  evaporated aluminum strips that define the electric field. At the bottom of field cage is the cathode that is  constructed from an aluminum honeycomb laminate, composed of two aluminum sheets bonded to an aluminum honeycomb core, providing a lightweight yet rigid structure. On the other end the boards were epoxied into an aluminum top perimeter which also served as the last electrode ring in the TPC. Together with the cathode bottom, the field cage proved to be a rigid lightweight structure. A Lexan ring containing o-rings was placed in-between the top perimeter piece and the top plate of the TPC. Screws with nylon washers, and collars, were used to mount the top perimeter, and the field cage, to the top plate; it also provided electrical isolation from the top plate. In this way the top plate could be removed and rotated with the field cage attached, without damaging any internal components.

@@

Figure~\ref{fig:FC_schematic} shows a schematic detailing the connections involved in the field cage walls. The field cage contains 50 inside copper strips and 49 outside copper strips. The strips are \SI{6}{\milli\metre} in width and spaced \SI{10}{\milli\metre} apart. Since the field cage was build with an array of PCB boards, connections from adjacent strips were connected all the way around until all the strips formed a continuous ring. The interfaces were the side pieces met were connected by G-10 corner pieces where conductive paint strips were painted on. The first inside strip is connected to the cathode, which is itself connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the first outside strip. The first outside strip is connected through a via to the second inside strip. These outside strips formed a small electric field which repelled electrons away from the PCB substrate between inside strips, preventing build up on the insulating material. This pattern repeats until the very last strip. A resistor chain, connected to the outside strips, creates a voltage divider in which each strip is separated by a constant difference voltage and a fixed distance, setting up a constant electric field. The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, which is the same as the other strip widths. The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Figure~\ref{fig:FC_schematic}. Allowing for a tunable resistor allowed for fine tuning of the electric field in the region between the top perimeter and the wire planes as will be discussed in Section~\ref{sec:wireplanes}.

@@

In the open configuration, the gating grid is transparent to electrons coming from the field cage volume and also allows for ions to move from the avalanche region into the TPC volume. In the closed configuration it is opaque to electrons and ions, when set to the right voltages. Typically the gating grid is always in the closed configuration, only opening it momentarily when the data acquisition trigger criteria is met to take data. By keeping it always closed the electrons which come from the un-reacted beam are blocked,  which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions, which would flood the field cage with space charge, distorting the electrons represnting the track paths. We open the gating grid for about \SI{11}{\micro\second} which is more than the time it takes for the electrons to drift one TPC volume. After this, we close the gating grid to prevent the back-flow of ions from the avalanche region from that event. Since ions move with a velocity much slower than that of electrons \cite{blumrol}, the ions only move several \si{\micro\metre} in the time the gate is open. This allows for electrons to pass through while preventing the back-flow of ions into the FC volume, reducing or even eliminating the space charge resulting from the avalanche region.

@@

The anode wires are made of very thin Gold plated Tungsten wires, about \SI{20}{\micro\metre} in diameter. They are biased to high voltages of about  \SI{e3}{\volt} which creates a very high electric field very close to the anode wire. As the electron approaches the anode wire, the field gets strong enough that the kinetic energy gained between collisions exceeds the ionization potential of the gas nocking out more electron-ion pairs in the gas. This happens over several collisions creating an avalanche effect amplifying the electron current.  The final number of electrons produced depends on the anode wire voltage and the gas properties. The absolute gas gain was not experimentally determined but it was calculated in a Garfield simulation. For the experimental data pertaining to this disertation, the anode wires were biased to two different voltages. We will refer to the voltage \SI{1460}{\volt} as the \emph{high voltage} and \SI{1214}{\volt} as the \emph{low voltage}. Two sections were biased with the lower voltage setting to minimize the effects of a leakage of secondary electrons around the end of the gating grid \cite{jon}.

@@

Care must be taken to choose the resister ($R_{TP2}$) that connects the field cage divider chain to the top plate. The end of the field cage nearest to the top plate is at the same potential as the aluminum top perimeter piece to which the field cage is glued. This top perimeter piece forms a horizontal surface that is an equipotential representing the last ring. The voltage difference between the average voltage on  the gating grid and the voltage of the top perimeter influences the electric field just below the gating grid. If $R_{TP2}$ is incorrectly chosen, the electric field near the gating grid will be different for electrons that pass near the field cage walls than it is for electrons in the center of the TPC,  causing distortions of the tracks in the TPC. This requires the value for $R_{TP2}$ to be chosen correctly. Since this simple point can be easily forgotten, we now discuss this choice in greater detail.

@@

The average energy loss in a material can be described by the Bethe-Bloch equation,
\begin{equation}\label{eq:bb}
\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (\ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2),
\end{equation}
where $N$ is the number density of electrons in the medium, $e$ the elementary charge, $mc^2$ is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle \cite{blumrol}. Yet there is a large variation in energy loss around this mean value. The statistical variation of energy loss in a material was described by Landau \cite{landau} and later better described by Shulek \cite{shulek} and Bichsel \cite{bichsel1}. In both approximations it is described by a most probable energy loss value, with a long, high-energy loss tail. The solid curve in Figure~\ref{fig:straggling} shows the energy loss distribution in Ar gas for a proton with momentum \SI{3.4}{\giga\eVperc} \cite{bichsel}. The dashed line is the distribution under the Landau assumptions. The mean energy loss $\langle\Delta\rangle$ is significantly shifted from the most probable value $\Delta_p$, due to the long high energy tail.  Because of this long tail, for a finite set of energy loss measurements along a given track, the mean value is a very unreliable. The most probable energy loss is the better observable which can be obtained either through fitting of the observed distribution or through the truncated mean method.

@@

where $f_r$ is the cut off fraction. In this thesis, a value of 0.7 was used and it was also used in the STAR TPC collaboration \cite{starsyst}.

@@

\section{Pad Response Function}
\label{sec:prf}
Each electron avalanche produces an two-dimensional image charge on the pad plane, as shown in the cartoon in Figure~\ref{fig:2DPRF}, where the projection of the charge distribution onto the $x$ and $z$ axis of this distribution are labeled as $\rho(x)$ and $\rho(z)$ respectively. If $\rho(x,z)$ represents the charge distribution on the pad-plane, the total charge observed a particular pad, $Q$, is expressed as,

@@

The \spirit TPC was designed to fit exactly into the dipole gap of the  dipole magnet at the end of the BigRIPS beam line. Figure~\ref{fig:experiment} shows a drawing of the \spirit TPC inside of the SAMURAI magnet chamber which was rotated to the $\ang{0}$ configuration. Typically the SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) dipole magnet is operated under vacuum as a large-acceptance multi-particle spectrometer for radioactive-beam experiments. This magnet can reach magnetic fields up to \SI{3}{\tesla} at the center of the pole gap and was operated at \SI{0.5}{\tesla} for these sets of experiments. The space between the magnetic pole faces is further complicated by large bolts which protrude from the pole faces. These bolts secure the vacuum chamber to the magnet which is not practically removable; though the inside of the magnet was not operated under vacuum. This required an extensive rail system and support frame to slowly slide in the TPC over the bolts, finally raising the TPC several \si{\centi\metre} to the final height.

@@

The height of the TPC was roughly aligned with a self-leveling laser system to match the center of target with the center of the beam line. Once the TPC was adjusted to the final location, the position of the TPC was measured in fine detail with a the VStars-N photogrametry system \cite{vstars}. Small highly reflective targets were placed all over the TPC both inside and out and pictures were taken with a calibrated lens and camera system. Using the commercial software provided the set of different camera perspectives was reconstructed into a 3-dimensional point cloud of all the targets. Since the magnet was also measured with the same system after installation, we can match the two systems to get the absolute position of the TPC, and its internal components, relative to the magnet frame. The position resolution of this type of system was estimated to be around \SI{200}{\micro\metre} for each coordinate, which is much more precise than needed or Show data.

The origin of the \spirit TPC was defined to be at the center of the TPC pad plane in the x-direction and at the edge of the last upstream pad. The center of the magnet is defined as the center of the magnet in x and z-directions and the middle of the dipole gap in the y-direction. The origin of the \spirit TPC in the magnet frame is (1.794,205.502,-580.526) for x,y,z-coordinates in units of \si{\milli\metre}. The error in each coordinate estimated to be (.2,.1,.4) \si{\milli\metre} respectively.

@@

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{TPCAux.png}
\caption{Exploded views of Kyoto and KATANA arrays.The Kyoto array elements are in tan and are located on the left an right side of the TPC. When they are in place, they are nested against the thin aluminum side windows of the TPC. The KATANA array is in green and is located adjacent to the downstream window of the TPC.}
\label{fig:aux}
\end{figure}

\subsection{Krakow KATANA Veto and Multiplicity Array}
%beam veto
%beam trigger optional
 Figure~\ref{fig:aux} also shows the Krakow KATANA array, which consists of 12 plastic scintillating bars mounted adjacent to the downstream wall of the TPC enclosure. Three of the 12 bars 1 mm in thickness and operated as a beam veto in the event the beam did not make a nuclear collision with the target; this occurred approximately 99\% of the time. The 9 other bars were 1 cm in thickness and provided additional information about the light particle multiplicity array similar to the Kyoto array. We found that a large number of events involved peripheral collisions and active target events in which the beam interactions with the counter gas. Active target events contributed to a large multiplicity in the KATANA but not to the Kyoto array. So multiplicity $M_{Kyoto} > 4$ in the Kyoto array combined with a veto on events with a forward going projectile remnant with Z>20 provided the most effective suppression of peripheral collisions and active target events. Thus the KATANA array was used in primarily the beam veto mode. This was accomplished by positioning the array so that the expected position of the beam exiting the TPC would be centered on the three thin paddles. The threshold of the veto paddles were set so that the charge of a particle passing through, $Z$, would veto any event that satisfied $Z > 20$, where the charge of the Sn beam is $Z=50$.

\subsection{Active Veto Array}
The beam was tuned by two sets of quadrupole  magnets, STQ 1 and STQ2 (as seen in Figure~\ref{fig:samuraiBeamLine}), so that the beam spot was focused on the TPC target location. Because of the inherent angular dispersion of the beam there still were incoming beam events which significantly deviated from the target location. These could be suppressed in the analyses by a gate on the reaction vertex; however, there remained a concern about the dead time such target frame events would contribute and how this would limit the amount of good data we could obtain.  To veto these type of events an active veto array was set at the entrance of the TPC consisting of four small scintillating bars arranged to be slightly larger than the target size. The threshold was set so that any beam particle which passed through any of the bars it would send a trigger signal to not trigger the system since the beam path would not be on target but on some other material inside the TPC. Minimizing these active veto rates also proved to be useful when tuning the beam so as to be centered on the target.

@@

There were several special trigger considerations incorporated into the trigger for the TPC. We required that the gating grid be opened quickly to not lose electrons from portions of the tracks that pass close to the gating grid. If the DAQ was not busy, we therefore generated a trigger once we had a start detector signal, a Kyoto multiplicity >4 signal, and neither a Katana Veto signal nor an active target signal. This was referred to as the Fast Trigger, and it triggered the opening of the gating grid, which took about 350 ns. Following the initial trigger, we had the possibility of a fast clear signal that could be initiated by a later beam particle traversing the TPC. If the last beam particle came within 4 us after the fast trigger, electrons from this beam particle could pass through the gating grid. Such data would be useless, so we closed the gating grid and stopped the data acquisition with a fast clear signal. Additional fast clears could also initiated if the Katana trigger box did not generate a trigger. Figure~\ref{fig:trigLogic} shows the logic of both of these triggers.

The Katana trigger box took some time to completely master, it did not take on the major trigger responsiblities at first. In later runs, most of the trigger functions were performed in the Katana trigger box. The master trigger for the DAQ was different for each primary beam. During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic, whereas in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics. In either case the differences in the trigger logic were very minor and they both behaved practically the same except for minor details on the gating-grid trigger \cite{jon}. Figure~\ref{fig:katanaLogic} summarizes the KATANA trigger box logic.

Figure~\ref{fig:dataTrigger1} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, KATANA multiplicity, no Veto, and no DAQ busy signal. The KATANA trigger, Kyoto Multiplicity, and Start Counter together triggered the DAQ.

>>>>

A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector. TPCs have been employed in gases, liquid, and even solid detectors \cite{starTPC,arTPC,eosTPC}. In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this Dissertation, along with the rest of the experimental setup.

@@

The pad plane is perpendicular to the electric field of the field cage and is divided into a two dimensional grid of pads. Two of the 3 coordinates are determined from the 2-dimensional charge distribution corresponding to the projection of the track onto the pad-plane. The third dimension comes from extrapolating the electrons back in time, utilizing the known constant drift velocity $v_d$. The distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the drift time information obtained from combining the time of the event trigger with arrival time of the signal from the  electrons in the electronics. The radius of curvature of the track on the pad plane is related to the magnetic rigidity, and therefore the momentum of each track. The energy loss deposited $\langle dE/dx\rangle$ is proportional to the image charge induced on the pads on the pad-plane. With these two observations, particles can be uniquely identified by their mass and charge since there is a unique correlation between the rigidity and $\langle dE/dx\rangle$  lines. This correlation allows the particle type to be identified as discussed later in Section~\ref{sec:pid}. In this chapter we will provide more details about the particle detection within the \spirit TPC  that is relevant to the reactions studied in this Dissertation.

@@

The SAMURAI Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) was developed to measure pions and other light charged particles produced in collisions of radioactive rare isotope beams with stable targets in fixed target experiments.  The TPC is built on an Aluminum angle iron skeleton with thin Aluminum sheet walls in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC. The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}. The dipole gap limited the vertical space of the TPC to around \SI{75}{\centi\metre}. More detail and specifications of the SAMURAI dipole magnet are given in \cite{samurai}.

The exploded drawing shown in Fig.~\ref{fig:tpcExplode} illustrates these major internal components of the  of the \spirit TPC. A target mechanism allowed for up to 5 fixed targets to be mounted on the target ladder situated just upstream of the field cage, where changing the targets could be performed outside of the magnet. Reaction products from nuclear collisions exit the target and enter the field cage of the TPC. The field cage contains the detector gas and defines the constant electric field. It is rigidly mounted to the large Aluminum top plate of the TPC enclosure. The field cage is electrically isolated from the top plate by a lexan top perimeter ring with o-rings to provide as gas seal. The pad plane and wire plane structures are  mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate. Several Aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a precise laser measurement \cite{jon}. Holes in the top plate allowed signals from the individual charge sensitive pads on the pad plane to be transmitted via cables to the electronics.

@@

\begin{comment}
\subsection{Enclosure}
The skeleton of the enclosure is composed of a rigid rectangular Aluminum angle-iron frame. To this frame, six rectangular sides are attached. The downstream window and the two side windows are constructed of a Aluminum window frames with a 1 mm  Aluminum sheet metal window. These thin sheet metal windows allow charged particles and neutrons to exit the TPC without significant energy loss or scattering. This allows for a trigger to be created by placing detectors outside of  the side windows and downstream window of the TPC. The enclosure itself is made to be gas tight with no leakage from the outside or from the interior of the field cage. This is to allow for the possibility to choosing one insulation gas outside of the field cage and a different counter gas inside the field cage. In the first campaign of experiments, however, we used the same gas in the field cage and enclosure volume.  
\end{comment}

@@

The field cage was constructed from several panels of two-layer printed circuit board (PCBs). The front of the field cage was made of two PCBs and each side was constructed from three PCBs. These boards had copper strips etched onto the inside and outside of the board which when all the boards were assembed, formed copper rings which were used to set up the uniform electric field, as will be described later. The array of circuit boards were glued and supported by Lexan pieces. We did not use the common PCB substrate material FR4, which contains a Bromine impregnated epoxy which can outgas Bromin and absorb the secondary electrons from the track, thereby degrading the signal \cite{tpcAging}. Instead, a halogen free material, Cryogenic-G10, was used for the board material.  The downstream wall mainly consists of a large, thin exit window. The window  is constructed of   \SI{10}{\micro\metre} Kapton upon which are  evaporated Aluminum strips that define the electric field. At the bottom of field cage is the cathode that is  constructed from an Aluminum honeycomb laminate, composed of two Aluminum sheets bonded to an Aluminum honeycomb core, providing a lightweight yet rigid structure. On the other end the boards were epoxied into an Aluminum top perimeter which also served as the last electrode ring in the TPC. Together with the cathode bottom, the field cage proved to be a rigid lightweight structure. A Lexan ring containing o-rings was placed in-between the top perimeter piece and the top plate of the TPC. Screws with nylon washers, and collars, were used to mount the top perimeter, and the field cage, to the top plate; it also provided electrical isolation from the top plate. In this way the top plate could be removed and rotated with the field cage attached, without damaging any internal components.

@@

Figure~\ref{fig:FC_schematic} shows a schematic detailing the connections involved in the field cage walls. The field cage contains 50 inside copper strips and 49 outside copper strips. The strips are \SI{6}{\milli\metre} in width and spaced \SI{10}{\milli\metre} apart. Since the field cage was built with an array of PCB boards, connections from adjacent strips were connected all the way around until all the strips formed a continuous ring. The interfaces where the side pieces met were connected by G10 corner pieces where conductive paint strips were painted on. The first inside strip is connected to the cathode, which is itself connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the first outside strip. The first outside strip is connected through a via to the second inside strip. These outside strips formed a small electric field which repelled electrons away from the PCB substrate between inside strips, preventing build up on the insulating material. This pattern repeats until the very last strip. A resistor chain, connected to the outside strips, creates a voltage divider in which each strip is separated by a constant difference voltage and a fixed distance, setting up a constant electric field. The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, which is the same as the other strip widths. The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Figure~\ref{fig:FC_schematic}. Allowing for a tunable resistor for fine tuning of the electric field in the region between the top perimeter and the wire planes as will be discussed in Section~\ref{sec:wireplanes}.

@@

In the open configuration, the gating grid is transparent to electrons coming from the field cage volume and also allows for ions to move from the avalanche region into the TPC volume. In the closed configuration it is opaque to electrons and ions, when set to the right voltages. Typically the gating grid is always in the closed configuration, only opening it momentarily when the data acquisition trigger criteria is met to take data. By keeping it always closed the electrons which come from the un-reacted beam are blocked,  which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions, flooding the field cage with space charge, distorting the electrons represnting the track paths. We open the gating grid for about \SI{11}{\micro\second} which is more than the time it takes for the electrons to drift 1 TPC volume. After this, we close the gating grid to prevent the back-flow of ions from the avalanche region from that event. Since ions move with a velocity much slower than that of electrons \cite{blumrol}, the ions only move several \si{\micro\metre} in the time the gate is open. This allows for electrons to pass through while preventing the back-flow of ions into the FC volume, reducing or even eliminating the space charge resulting from the avalanche region.

@@

The anode wires are made of very thin Gold plated Tungsten wires, about \SI{20}{\micro\metre} in diameter. They are biased to high voltages of about  \SI{e3}{\volt} which creates a very high electric field very close to the anode wire. As the electron approaches the anode wire, the field gets strong enough that the kinetic energy gained between collisions exceeds the ionization potential of the gas knocking out more electron-ion pairs in the gas. This happens over succesive  collisions creating an avalanche effect amplifying the electron current.  The final number of electrons produced depends on the anode wire voltage and the gas properties. The absolute gas gain was not experimentally determined but it was calculated in a Garfield simulation. For the experimental data pertaining to this disertation, the anode wires were biased to two different voltages. We will refer to the voltage \SI{1460}{\volt} as the \emph{high voltage} and \SI{1214}{\volt} as the \emph{low voltage}. Two sections of the anode wire plane, were biased with the lower voltage setting to minimize the effects of a leakage of secondary electrons around the end of the gating grid \cite{jon}.

@@

Care must be taken to choose the resister ($R_{TP2}$) that connects the field cage divider chain to the top plate. The end of the field cage nearest to the top plate is at the same potential as the Aluminum top perimeter piece to which the field cage is glued. This top perimeter piece forms a horizontal surface that is an equipotential representing the last ring. The voltage difference between the average voltage on  the gating grid and the voltage of the top perimeter influences the electric field just below the gating grid. If $R_{TP2}$ is incorrectly chosen, the electric field near the gating grid will be different for electrons that pass near the field cage walls than it is for electrons in the center of the TPC,  causing distortions of the tracks in the TPC. This requires the value for $R_{TP2}$ to be chosen correctly.

@@

The average energy loss in a material can be described by the Bethe-Bloch equation,
\begin{equation}\label{eq:bb}
\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (\ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2),
\end{equation}
where $N$ is the number density of electrons in the medium, $e$, the elementary charge, $mc^2$, is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle \cite{blumrol}. Yet there is a large variation in energy loss around this mean value. The statistical variation of energy loss in a material was described by Landau \cite{landau} and later better described by Shulek \cite{shulek} and Bichsel \cite{bichsel1}. In both approximations it is described by a most probable energy loss value, with a long, high-energy loss tail. The solid curve in Figure~\ref{fig:straggling} shows the energy loss distribution in Ar gas for a proton with momentum \SI{3.4}{\giga\eVperc} \cite{bichsel}. The dashed line is the distribution under the Landau assumptions. The mean energy loss $\langle\Delta\rangle$ is significantly shifted from the most probable value $\Delta_p$, due to the long high energy tail.  Because of this long tail, for a finite set of energy loss measurements along a given track, the mean value is very unreliable. The most probable energy loss is the better observable which can be obtained either through fitting of the observed distribution or through the truncated mean method.

@@

where $f_r$ is the cut off fraction. In this Dissertation, a value of 0.7 was used. This was the same value used in the STAR TPC collaboration \cite{starsyst}.

@@

\section{Pad Response Function}
\label{sec:prf}
Each electron avalanche produces a two-dimensional image charge on the pad plane, as shown in the cartoon in Figure~\ref{fig:2DPRF}, where the projection of the charge distribution onto the $x$ and $z$ axis of this distribution are labeled as $\rho(x)$ and $\rho(z)$ respectively. If $\rho(x,z)$ represents the charge distribution on the pad-plane, the total charge observed a particular pad, $Q$, is expressed as,

@@

The \spirit TPC was designed to fit into the dipole gap of the  dipole magnet at the end of the BigRIPS beam line. Figure~\ref{fig:experiment} shows a drawing of the \spirit TPC inside of the SAMURAI magnet chamber which was rotated to the $\ang{0}$ configuration. Typically the SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) dipole magnet is operated under vacuum as a large-acceptance multi-particle spectrometer for radioactive-beam experiments. This magnet can reach magnetic fields up to \SI{3}{\tesla} at the center of the pole gap and was operated at \SI{0.5}{\tesla} for these sets of experiments. The space between the magnetic pole faces is further complicated by large bolts which protrude from the pole faces. These bolts secure the vacuum chamber to the magnet which is not practically removable; though the inside of the magnet was not operated under vacuum. This required an extensive rail system and support frame to slowly slide in the TPC over the bolts, finally raising the TPC several \si{\centi\metre} to the final height.

@@

The height of the TPC was roughly aligned with a self-leveling laser system to match the center of target with the center of the beam line. Once the TPC was adjusted to the final location, the position of the TPC was measured in fine detail with a the VStars-N photogrametry system \cite{vstars}. Small highly reflective targets were placed all over the TPC both inside and out and pictures were taken with a calibrated lens and camera system. Using the commercial software provided the set of different camera perspectives was reconstructed into a 3-dimensional point cloud of all the targets. Since the magnet was also measured with the same system after installation, we can match the two systems to get the absolute position of the TPC, and its internal components, relative to the magnet frame. The position resolution of this type of system was estimated to be around \SI{200}{\micro\metre} for each coordinate.

The origin of the \spirit TPC was defined to be at the center of the TPC pad plane in the x-direction and at the edge of the first upstream pad. The center of the magnet is defined as the center of the magnet in x and z-directions and the middle of the dipole gap in the y-direction. The origin of the \spirit TPC in the magnet frame is (1.8,205.5,-580.5) for x,y,z-coordinates in units of \si{\milli\metre}. The error in each coordinate estimated to be (0.2,0.1,0.4) \si{\milli\metre} respectively.

@@

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{TPCAux.png}
\caption{Exploded views of Kyoto and KATANA arrays.The Kyoto array elements are in tan and are located on the left an right side of the TPC. When they are in place, they are nested against the thin Aluminum side windows of the TPC. The KATANA array is in green and is located adjacent to the downstream window of the TPC.}
\label{fig:aux}
\end{figure}

\subsection{KATANA Veto and Multiplicity Array}
%beam veto
%beam trigger optional
 Figure~\ref{fig:aux} also shows the Krakow KATANA array, which consists of 12 plastic scintillating bars mounted adjacent to the downstream wall of the TPC enclosure. Three of the 12 bars, which are 1 mm in thickness, and operated as a beam veto in the event the beam did not make a nuclear collision with the target; this occurred approximately 99\% of the time. The 9 other bars were 1 cm in thickness and provided additional information about the light particle multiplicity array similar to the Kyoto array. We found that a large number of events involved peripheral collisions and active target events in which the beam interactions with the counter gas. Active target events contributed to a large multiplicity in the KATANA but not to the Kyoto array. So multiplicity $M_{Kyoto} > 4$ in the Kyoto array combined with a veto on events with a forward going projectile remnant with Z>20 provided the most effective suppression of peripheral collisions and active target events. Thus the KATANA array was used in primarily the beam veto mode. This was accomplished by positioning the array so that the expected position of the beam exiting the TPC would be centered on the three thin paddles. The threshold of the veto paddles were set so that the charge of a particle passing through, $Z$, would veto any event that satisfied $Z > 20$, where the charge of the Sn beam is $Z=50$.

\subsection{Active Veto Array}
The beam was tuned by two sets of quadrupole  magnets, STQ 1 and STQ2 (as seen in Figure~\ref{fig:samuraiBeamLine}), so that the beam spot was focused on the TPC target location. Because of the inherent angular dispersion of the beam, some events significantly deviated from the target location. These could be suppressed in the analyses by a gate on the reaction vertex; however, there remained a concern about the dead time such target frame events would contribute and how this would limit the amount of good data we could obtain.  To veto these type of events an active veto array was set at the entrance of the TPC consisting of four small scintillating bars arranged to be slightly larger than the target size. The threshold was set so that any beam particle which passed through any of the bars it would send a signal not to trigger the system since the beam path would not be on target but on some other material inside the TPC. Minimizing these active veto rates also proved to be useful when tuning the beam to be more centered on the target.

@@

There were several special trigger considerations incorporated into the trigger for the TPC. We required that the gating grid be opened quickly to not lose electrons from portions of the tracks that pass close to the gating grid. If the DAQ was not busy, we therefore generated a trigger once we had a start detector signal, a Kyoto multiplicity >4 signal, and neither a Katana Veto signal nor an active target signal. This was referred to as the Fast Trigger, and it triggered the opening of the gating grid, which took about 350 ns. Following the initial trigger, we had the possibility of a fast clear signal that could be initiated by a later beam particle traversing the TPC. If the last beam particle came within 4 us after the fast trigger, electrons from this beam particle could pass through the gating grid. Such data would be useless, so we closed the gating grid and stopped the data acquisition with a fast clear signal. Additional fast clears  also was initiated if the Katana trigger box did not generate a trigger. Figure~\ref{fig:trigLogic} shows the logic of both of these triggers.

The Katana trigger box initially did not take on the major trigger responsiblities. In later runs, most of the trigger functions were performed in the Katana trigger box. The master trigger for the DAQ was different for each primary beam. During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic, whereas in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics. In either case the differences in the trigger logic were very minor and they both behaved practically the same except for minor details on the gating-grid trigger \cite{jon}. Figure~\ref{fig:katanaLogic} summarizes the KATANA trigger box logic.

Figure~\ref{fig:dataTrigger1} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output required a Start Counter, KATANA multiplicity, no Veto, and no DAQ busy signal. The KATANA trigger, Kyoto Multiplicity, and Start Counter together triggered the DAQ.

 Figure~\ref{fig:dataTrigger2} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output required a Start Counter, Kyoto or KATANA multiplicity, no Veto, and no DAQ busy signal. Here the KATANA trigger and the SC SUM together triggered the DAQ.
